---
title: "Tutorial"
author: "Brittany Maldonado"
date: "11/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#First thing - how to install the package. Syntax:

```{r}
devtools::install_github("brittanymaldonado/maldonadopackage")
```

#Next - reminder to load package. 

```{r}
library(maldonadopackage)
```

## Formatting dates and creating "Year" column

#Do you have a long-term dataset with many observations? Would you like to format your data in a way that allows easy visualization of year-to-year trends? Then this is the function for you! The dataset you input only needs to have a "Date" column in the following format: MM/DD/YYYY. This function is ideal for long-term datasets with observations across years or many observations per month. What do you get out of it? A column, "Year," that pulls the year for each observation out of the date and a column, "Month" that pulls the month for each observation out of the date. 


```{r}
library(tidyverse)
library(assertthat)
library(lubridate)
agp <- read.csv(file = "../inst/extdata/agpdata.csv")
yearify(agp)
```
